---
title: "Project title"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false

library(tidyverse)
```

# Data 1

## Introduction and data

-   Identify the source of the data.

Our data is sourced from the 2018 Central Park Squirrel Census, compiled by NYC Open Data.

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

The data was collected in October 2018 by hundreds of volunteers and key NYC entities who tallied the squirrels in Central Park, New York on behalf of The Squirrel Census (the original data curator).

-   Write a brief description of the observations.

The observations include 3,023 squirrel sightings, unique squirrel IDs, dates spotted, location, age, primary and secondary fur color, elevation, squirrel activities when spotted (running, chasing, climbing, foraging, eating, or other), sounds, tail movement, and interactions between squirrels and with humans.

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

-   Identify the types of variables in your research question. Categorical? Quantitative?

## Literature

-   Find one published credible article on the topic you are interested in researching.

-   Provide a one paragraph summary about the article.

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

## Glimpse of data

```{r}
#| label: load-data-1

# add code here
```

# Data 2

## Introduction and data

-   Identify the source of the data.

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

-   Write a brief description of the observations.

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

-   Identify the types of variables in your research question. Categorical? Quantitative?

## Literature

-   Find one published credible article on the topic you are interested in researching.

-   Provide a one paragraph summary about the article.

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

## Glimpse of data

```{r}
#| label: load-data-2

# add code here
```

# Data 3

## Introduction and data

-   Identify the source of the data.

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

-   Write a brief description of the observations.

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

-   Identify the types of variables in your research question. Categorical? Quantitative?

## Literature

-   Find one published credible article on the topic you are interested in researching.

-   Provide a one paragraph summary about the article.

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

## Glimpse of data

```{r}
#| label: load-data-3

# add code here
```
