---
title: "Project title"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false

library(tidyverse)
```

# Data 1

## Introduction and data

-   Identify the source of the data.

**Our data is sourced from the 2018 Central Park Squirrel Census, compiled by NYC Open Data.**

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

**The data was collected in October 2018 by hundreds of volunteers and key NYC entities who tallied the squirrels in Central Park, New York on behalf of The Squirrel Census (the original data curator).**

-   Write a brief description of the observations.

**The observations include 3,023 squirrel sightings, unique squirrel IDs, dates spotted, location, age, primary and secondary fur color, elevation, squirrel activities when spotted (running, chasing, climbing, foraging, eating, or other), sounds, tail movement, and interactions between squirrels and with humans.**

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

**Research question: What is the relationship between squirrel location, squirrel climbing, and whether the squirrel runs from human spotters?**

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

**Our hypothesis is that squirrels spotted on a ground plane will run from human spotters more than squirrels above ground, given the spotter is on their same elevation and thus constitutes a greater threat. Moreover, we hypothesize that squirrels climbing will run from human spotters more than squirrels above ground and more than squirrels spotted on a ground plane, given such squirrels might climb to escape the human.**

-   Identify the types of variables in your research question. Categorical? Quantitative?

**All variables are categorical variables - climbing, location, runs from.**

## Literature

-   Find one published credible article on the topic you are interested in researching.

**https://www.jstor.org/stable/3800817?seq=8**

**https://www.jstor.org/stable/3797645**

-   Provide a one paragraph summary about the article.

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

## Glimpse of data

```{r}
#| label: load-data-1

glimpse(Squirrels)
```

# Data 2

## Introduction and data

-   Identify the source of the data.

This data is sourced from the Division I college basketball seasons from the years of 2013-2019. It was found in the open source data site, Kaggle.

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

The data was scraped from <http://barttorvik.com/trank.php#>. This data was cleaned up in 2021, in which the COVID seasons were not included in the dataset.

-   Write a brief description of the observations.

The dataset has various variables that characterize different Division I basketball teams--- such as the university they represent and the conference in which they belong. Some of the observations within the dataset include the number of games played, the number of games won, power rating, as well as the stats of the team in a particular season.

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

**Research question:** Which Division 1 basketball conference has the highest success rate from the 2013-2019 season? (number of game wins, offensive rebound rates, total points, etc.). And which best factors (other than wins)(offensive/ defensive efficiency, 2pt and 3pt field goal success, steal rate, turnover rate, etc) best predict regular season success (win percentage)(from 2013 to 2019)? And does this also predict (better predict?/ worse predict) post season success?

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

Our research topic will explore the different facets that are considered in possibly predicting the success of college DI basketball teams in 2013-2019. We hypothesize that the best predictors for regular season success (win percentage) will be offensive and defensive efficiency and it will also predict postseason success just as well.

-   Identify the types of variables in your research question. Categorical? Quantitative?

Conference (categorical)

Wins(quantitative)

Losses(quantitative)

Adjusted offensive rating(quantitative)

Adjusted defensive rating(quantitative)

3pt success rate (allowed and committed)(quantitative)

2pt success rate (allowed and committed)(quantitative)

Field goal success rate(quantitative)

Effective field goal percentage (allowed and committed)(quantitative)

Rebound and turnover rates (allowed and committed)(quantitative)

How to calculate success

Post season- How far they made it in March madness (not in tournament, round of 64, round of 32, sweet 16, elite 8, final 4, runner up, champions) (categorical)

Regular season - Seed for the tournament (categorical) (easily changeable to quantitative ) and Win percentage (quantitative)

## Literature

-   Find one published credible article on the topic you are interested in researching.

    -   <https://www.ncaa.com/news/basketball-men/article/2018-10-05/best-conferences-ncaa-tournament-play-1985>

-   Provide a one paragraph summary about the article.

    -   This article looks at the success of conferences since the NCAA tournament was made into a 64 team format (now 68). With the ACC being largely in control statistically, they are the dominant post season  conference (which is argued to be the best team predictor since in the tournament teams from the same conference rarely play each other). The ACC also has a winning record against 29 of the 32 conferences. The ACC is also the league's only double digit champion (with 10 titles since 1985) and has the highest win percentage. (\*This article is also reputable as it uses data directly from the NCAA and it written and researched by the NCAA)

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

    -   In our research question we wonder why this is the pattern. Our data should show this same pattern, but it should also suggest a predictor for this behavior as well. Why do some teams perform at a higher success and what are the variables that predict success? 

## Glimpse of data

```{r}
#| label: load-data-2

glimpse(cbb)
```

# Data 3

## Introduction and data

-   Identify the source of the data.


-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

-   Write a brief description of the observations.

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

-   Identify the types of variables in your research question. Categorical? Quantitative?

## Literature

-   Find one published credible article on the topic you are interested in researching.

-   Provide a one paragraph summary about the article.

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

## Glimpse of data

```{r}
#| label: load-data-3

glimpse(the_counted_2016)

```
